\section{Project Organization}
\label{section:project_organization}

All work in this project was realized within the TUM Data Innovation Lab in collaboration with the Chair of 
Aerodynamics and Fluid Mechanics. Special thanks go to our Scientific Lead PhD candidate Ludger P\"ahler, 
as well as our Project Lead and Co-Mentor Dr. Ricardo Acevedo Cabra.

Our project was structured in 4 phases:

\begin{enumerate}
    \item \textit{Literature review (4 weeks)}: We read through important papers in the topics of our project such 
    as Graph Learning, Graph Neural Networks, Graph Sampling, Message Passing for Quantum Chemistry, 
    and Scalable Machine Learning; and discussed them weekly.
    \item \textit{Baseline (3 weeks)}: After getting into the topic theoretically, we started looking at the OCP 
    project and elaborating the current state of the project to see what baseline we are comparing against.
    \item \textit{Implementation (4 weeks)}: After defining the baseline, we integrated and tested all DeepSpeed 
    stages in the OCP project during our implementation phase.
    \item \textit{Evaluation and Final report (3 weeks)}: In the final weeks of the project, we conducted baselines 
    and DeepSpeed experiments and recorded our observations and analysis. In addition, we focused on the final 
    report in this phase.
\end{enumerate}

Our team met twice each week on Monday and Friday to discuss progress and define next steps. Most of 
the time, individual team members met two to three additional times during the week to discuss specific 
tasks in their area of activity. For organizational purposes, team and personal conversations, we used 
a private Slack channel.

For version management of our code we created a GitHub organization\footnote{\url{https://github.com/TUM-DI-Lab-Graph-Scaling}} 
where we placed multiple projects like our own fork of the OCP project, evaluation scripts and also 
the source files for the final report. We tracked problems and errors we encountered as issues in the 
corresponding projects and worked with a branch structure to avoid disrupting the work of our teammates.