\section{Evaluation}
\label{section:evaluation}

\subsection{Datasets}

The OC20 dataset \cite{Chanussot_2021} contains training and evaluation data of roughly 
1.2M DFT relaxations using a range of different materials, surfaces and adsorbates. 
Jointly with the dataset, three different challenges for the community were published:
IS2RE (Initial Structure to Relaxed Energy), S2EF (Structure to Energy and Forces) and
IS2RS (Initial Structure to Relaxed Structure). In our work, we focused on the first two.

\subsubsection{Initial Structure to Relaxed Energy (IS2RE)}

The Initial Structure to Relaxed Energy (IS2RE) task takes an initial atomic 
structure as input and predicts the energy after relaxation of the structure 
\cite*{Chanussot_2021, https://doi.org/10.48550/arxiv.2206.08917}. 


OC20 provides three dataset scaling factors for the IS2RE task: 10k, 100k and 
all structures + relaxations. Each dataset is a subset of the next higher-scaled 
dataset.

\subsubsection{Structure to Energy and Forces (S2EF)}

The Structure to Energy and Forces (S2EF) task takes an atomic structure and 
predicts energy of the structure and per-atom forces 
\cite*{Chanussot_2021, https://doi.org/10.48550/arxiv.2206.08917}. Four metrics 
are tracked during the S2EF task: Mean Absolute Error for the energy and forces 
(MAE), Force Cosine similarity, and energy as well as forces within a threshold 
(EFwT). 

OC20 provides three dataset scaling factors for the S2EF task: 200k, 2M, 20M,  
and all. The scaling factor determines the number of atomic structures and 
pre-computed energy and per-atom forces with DFT. The input structures are 
provided as compressed trajectory files, so we had to locally uncrompress the 
files and embedded the structures into Lightning Memory-Mapped Databases (LMBDs). 
Similiar to the datasets in IS2RE, each dataset is a subset of the next 
higher-scaled dataset. 